---
title: "Individual Project"
author: "Saeun Park"
output: html_document
---

# The World Happiness Report 2024

## 0. Load Libraries and Dataset

```{r}
rm(list=ls())
```

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)
library(reshape2)
library(tidyverse) # Clustering
library(conflicted) 
library(factoextra) 
```

```{r,eval=TRUE}
rm(list=ls())
whr = read.csv('/Users/saeun/Desktop/Courses/Modern computer-based methods of statistics/individual_project/WHR2024.csv')

```

## 1. Data Structure

```{r,eval=TRUE}
head(whr)
dim(whr) # (143, 11)
str(whr)
summary(whr)
```

## 2. Data Preprocessing

### 2.1. Rename Columns

```{r,eval=TRUE}
colnames(whr) = c('Country', 'Score', 'upperwhisker', 'lowerwhisker',
                  'LogGDP', 'SocialSupport', 'LifeExpectancy', 
                  'Freedom', 'Generosity', 'Corruption', 'DystopiaResidual')
```

### 2.2. Remove Missing Values

Removing 'Bahrain', 'Tajikistan', 'State of Palestine'

```{r}
colSums(is.na(whr))
whr[apply(whr, 1, function(x) any(is.na(x))), ] # Identifying Rows with Missing Values
whr = whr[!apply(whr, 1, function(x) any(is.na(x))), ] # Remove the Rows
dim(whr) # (140, 11)
```

### 2.3. Add Region and SubRegion Columns to the Dataset

<https://www.kaggle.com/datasets/andradaolteanu/country-mapping-iso-continent-region>

```{r}
# Mapping Using External Data
cont = read.csv('/Users/saeun/Desktop/Courses/Modern computer-based methods of statistics/individual_project/continents2.csv')
head(cont)
cont = cont[,c('name', 'region', 'sub.region')]
colnames(cont) = c('Country', 'Region', 'SubRegion')
whr_new = merge(whr, cont, by='Country', all.x=TRUE)

# Manually Mapping 10 Unmapped Rows
whr_new[is.na(whr_new$Region),]['Country']
map = data.frame(
  Country = c("Bosnia and Herzegovina", "Congo (Brazzaville)", "Congo (Kinshasa)", 
              "Czechia", "Hong Kong S.A.R. of China", "Ivory Coast", 
              "Kosovo", "North Macedonia", "Taiwan Province of China", "Turkiye"),
  Region = c("Europe", "Africa", "Africa", 
             "Europe", "Asia", "Africa", 
             "Europe", "Europe", "Asia", "Asia"),
  SubRegion = c("Southern Europe", "Sub-Saharan Africa", "Sub-Saharan Africa", 
                "Eastern Europe", "Eastern Asia", "Sub-Saharan Africa", 
                "Southern Europe", "Southern Europe", "Eastern Asia", "Western Asia")
)
whr_new = merge(whr_new, map, by='Country', all.x=TRUE)

whr_new$Region = ifelse(is.na(whr_new$Region.x), whr_new$Region.y, whr_new$Region.x)
whr_new$SubRegion = ifelse(is.na(whr_new$SubRegion.x), whr_new$SubRegion.y, whr_new$SubRegion.x)

whr_new = whr_new %>%
  dplyr::select(-Region.x, -SubRegion.x, -Region.y, -SubRegion.y)

# Verifying the Results
sum(is.na(whr_new['Region']))
```

## 3. Data Visualization

### 3.1. Setting

num_var : Storing Only Numeric Variables

num_data : Storing Numeric Variable Data

```{r}
num_var = setdiff(colnames(whr_new), c("Country", 'upperwhisker', 'lowerwhisker', "Region", "SubRegion"))
num_data = whr_new[,num_var]
```

### 3.2. Boxplots for Each Variable

```{r}
plot_list = list()
for (var in num_var) {
  p = boxplot(num_data[[var]], main = paste("Boxplot of", var), col = "lightblue", outline = TRUE)
  
  plot_list[[var]] = p
}
```

### 3.3. Density Distribution for Each Variable

```{r}
plot_list1 = list()

for (var in num_var) {
  p = ggplot(whr_new, aes_string(x = var)) +
    geom_density(fill = "blue", alpha = 0.3) +
    labs(title = paste("Distribution of", var), x = var, y = "Density") +
    theme_minimal()
  
  plot_list1[[var]] = p
}
print(plot_list1)
#grid.arrange(grobs = plot_list1, ncol = 2)
```

### 3.4. Correlation Matrix

```{r}
corr_matrix = cor(num_data, use = "complete.obs")
melt = melt(corr_matrix) # melt dimension

ggplot(data = melt, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  geom_text(aes(label = round(value, 2)), size = 3, color = "black") + 
  scale_fill_gradient(low = "white", high = "darkgreen") +
  labs(title = "Correlation Matrix") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

### 3.5. Relationship between LifeExpectancy And LogGDP

```{r}
ggplot(whr_new, aes(x = LifeExpectancy, y = LogGDP)) +
  geom_point(color = "red", alpha = 0.7) +
  geom_smooth(method = "lm", color = "darkred") +
  labs(title = "Relationship between A and B", 
       x = "LifeExpectancy", y = "LogGDP")
```

### 3.6. Relationship between SocialSupport And Score

```{r}
ggplot(whr_new, aes(x = SocialSupport, y = Score)) +
  geom_point(color = "blue", alpha = 0.7) +
  geom_smooth(method = "lm", color = "darkblue") +
  labs(title = "Relationship between A and B", 
       x = "SocialSupport", y = "Score")
```

### 3.7. Barplots of the Interesting Countries for Each Variable

```{r}
southkorea_neighbors = c("South Korea", "Japan", "China")
hungary_neighbors = c("Hungary", "Slovakia", "Austria", "Czechia", 
                       "Slovenia", "Croatia", "Serbia", "Romania")
highscore_countries = c("Finland", "Denmark", "Iceland", "Sweden")

interesting_countries = c(southkorea_neighbors, hungary_neighbors, highscore_countries)

plot_list2 = list()
for (var in num_var) {
  filtered_data = whr_new %>%
    dplyr::filter(Country %in% interesting_countries)
  
  filtered_data = filtered_data %>%
    mutate(Group = case_when(
      Country %in% southkorea_neighbors ~ "South Korea & Neighbors",
      Country %in% hungary_neighbors ~ "Hungary & Neighbors",
      Country %in% highscore_countries ~ "High Score Countries"
    ))
  
  p = ggplot(filtered_data, aes(x = reorder(Country, .data[[var]]), y = .data[[var]], fill = Group)) +
    geom_bar(stat = "identity", position = "dodge") +
    geom_text(aes(label = SubRegion), position = position_dodge(width = 0.9), size = 2) +
    coord_flip() +
    labs(title = paste("Barplot of", var), x = "Country", y = var) +
    scale_fill_manual(values = c("South Korea & Neighbors" = "lightblue", 
                                 "Hungary & Neighbors" = "orange", 
                                 "High Score Countries" = "lightgreen")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      axis.text.y = element_text(size = 7),
      legend.position = "bottom"
    )
  
  plot_list2[[var]] = p
}
plot_list2

```

## 4. Data Analysis

### 4.1. Clustering Analysis

#### 4.1.1 Determining the Number of Clusters

Using Elbow Method (k=5)

```{r}
num_var = setdiff(num_var, "Score")
df = whr_new[,num_var]
# Scaling Data
df = scale(df)

# Determine the optimal number of clusters using the Elbow Method
wss = numeric(30) # Within-cluster Sum of Square
for (k in 1:30) {
  model = kmeans(df, centers = k, nstart = 25)
  wss[k] = model$tot.withinss
}

# Visualize the Elbow Plot
plot(1:30, wss, type = "b", pch = 19, xlab = "Number of Clusters (K)", ylab = "Total Within-SS")
```

Using Hierarchial Clustering (k=9)

```{r}
dist = dist(df)
hc = hclust(dist, method = "ward.D2")

# Visualize Dendrogram
plot(hc, labels = FALSE, main = "Dendrogram", xlab = "", sub = "")
clusters_hc = cutree(hc, k = 10)
```

```{r}
k = 10

set.seed(123)
kmeans_result = kmeans(df, centers = k, nstart = 25)

# Add clustering results to the whr_new data
whr_new$Cluster = as.factor(kmeans_result$cluster)
```

#### 4.1.2. Plotting the Result of Clustering Analysis

```{r}
all(rownames(df) == rownames(whr_new)) #True

fviz_cluster(kmeans_result, data = df, geom = "point", ellipse.type = "norm", 
             ggtheme = theme_minimal(), main = "K-Means Clustering Results") + 
  geom_text(aes(label = whr_new$Country), size = 2.5, vjust = -1)
```

```{r}
# Average Score by Cluster
cluster_avg_score = whr_new %>%
  group_by(Cluster) %>%
  summarise(avg_score = mean(Score, na.rm = TRUE))

ggplot(cluster_avg_score, aes(x = as.factor(Cluster), y = avg_score, fill = as.factor(Cluster))) +
  geom_bar(stat = "identity") +
  labs(
    title = "Average Score by Cluster",
    x = "Cluster",
    y = "Average Score"
  ) +
  scale_fill_brewer(palette = "Set3") +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

```{r}
## Boxplots for each variable grouped by clusters
boxplots = lapply(num_var, function(var) {
  ggplot(whr_new, aes(x = as.factor(Cluster), y = .data[[var]], fill = as.factor(Cluster))) +
    geom_boxplot() +
    labs(title = paste("Boxplot of", var), x = "Cluster", y = var) +
    theme_minimal() +
    theme(legend.position = "none")
})
# grid.arrange(grobs = boxplots, ncol = 3) 
print(boxplots)
```

```{r}
# Computing Averages for Each Cluster
summary = whr_new %>%
  group_by(Cluster) %>%
  summarise(across(c(LogGDP, SocialSupport, LifeExpectancy, Freedom, Generosity, Corruption, DystopiaResidual), mean, na.rm = TRUE))

print(summary)

# Wide format -> Long format
long_summary = summary %>%
  pivot_longer(cols = -Cluster, names_to = "Variable", values_to = "Mean")

# Visualization 
ggplot(long_summary, aes(x = Variable, y = Mean, fill = as.factor(Cluster))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Mean of Variables by Cluster", x = "Variable", y = "Mean") +
  theme_minimal()
```

```{r}
# Number of Regions by Cluster
region_summary = whr_new %>%
  group_by(Cluster, Region) %>%
  summarise(Count = n(), .groups = "drop") %>%
  arrange(Cluster, desc(Count))

# Visualization
ggplot(region_summary, aes(x = as.factor(Cluster), y = Count, fill = Region)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_brewer(palette = "Set1") + 
  labs(title = "Number of Regions by Cluster", x = "Cluster", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "top")
```

#### 4.1.3. Interesting Countries

```{r}
interesting_df = whr_new %>%
  dplyr::filter(Country %in% interesting_countries)
interesting_df[c('Country', 'Cluster')]
```

```{r}
ggplot(whr_new, aes(x = LogGDP, y = Score, color = Cluster)) +
  geom_point(alpha = 0.6) +
  geom_point(data = interesting_df, aes(x = LogGDP, y = Score), color = "black", size = 3) +
  ggtitle("World Happiness by Log GDP with Interesting Countries") +
  theme_minimal()
```

```{r}
# Visualization of Cluster Distribution by Region
region_distribution = interesting_df %>%
  group_by(Region, Cluster, Country) %>%
  summarise(count = n(), .groups = 'drop')

ggplot(region_distribution, aes(x = Region, fill = as.factor(Cluster), y = count)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = Country), position = position_stack(vjust = 0.5), size = 3, check_overlap = TRUE) +
  ggtitle("Cluster Distribution by Region") +
  theme_minimal()

# Visualization of Cluster Distribution by SubRegion
subregion_distribution = interesting_df %>%
  group_by(SubRegion, Cluster, Country) %>%
  summarise(count = n(), .groups = 'drop')

ggplot(subregion_distribution, aes(x = SubRegion, fill = as.factor(Cluster), y = count)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = Country), position = position_stack(vjust = 0.5), size = 3, check_overlap = TRUE) +
  ggtitle("Cluster Distribution by SubRegion") +
  theme_minimal()

```

### 4.2. Principal Component Analysis

```{r}
df_pca = whr_new %>%
  select(-Cluster, -Score, -upperwhisker, -lowerwhisker) %>%
  select_if(is.numeric)  # select only numeric variables

pca_res = prcomp(df_pca, scale = TRUE)

# Scree Plot
fviz_eig(pca_res, addlabels = TRUE, geom = "bar", bar_fill = "skyblue")

# Contributions of Variables
pca_res$rotation
fviz_pca_var(pca_res, col.var = "contrib", gradient.cols = c("blue", "green", "red"))
```

### 4.3. Logistic Regression

```{r}
# Extraction of Principal Components (PC1, PC2, PC3)
pca_scores = data.frame(pca_res$x[, 1:3])
colnames(pca_scores) = paste0("PC", 1:3)

logit_data = cbind(pca_scores, Cluster = whr_new$Cluster)

# Logistic Regression Model
library(nnet)
logit = multinom(Cluster ~., data = logit_data)
summary(logit)
```

```{r}

z_scores = summary(logit)$coefficients / summary(logit)$standard.errors
p_values = (1 - pnorm(abs(z_scores), 0, 1)) * 2
print(p_values)
```
